---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

### Introduction and Data

[INTRODUCTION]{.underline}

Figure skating is a sport in which athletes complete programs full of jumps, spins, and footwork while making sure to perform to their music. Scoring for figure skating consists of two components, the technical score and the artistic score. And it's this scoring that is the source of much of the controversy in the figure skating world.

In recent years, women's figure skating has experienced a "quad revolution" of sorts, with female skaters beginning to perform quadruple jumps with a higher frequency. This has been met with many questions, from fans, coaches, and authorities alike. Are quadruple jumps -- where the skater rotate four revolutions in the air -- valued too heavily? Are they leading to degradation in the performance of other elements?

One of the skaters that helped bring this revolution about is Alexandra Trusova, a Russian figure skater whose national and international competitive career spanned from 2014 to 2022. She was the first to perform several types of quadruple jumps in competition, and her earlier competitive years are well documented. Thus, she is a prime subject for analysis.

[THE DATA]{.underline}

Our data is collected from the competition records of the International Skating Union, Tallinn Trophy, and the Russian Skating Federation. The International Skating Union (ISU) is the governing body of figure skating competition around the world. ISU-sanctioned and many other competitions have websites containing records of the detailed results of the competition. Each event has an attached pdf file, recording every skater's detailed scoring breakdown, including the judges' evaluation of their jumps. After obtaining Alexandra Trusova's competitive history, the corresponding data for each of her jumps completed in national championships or international competition was pulled into a dataset, accompanied by relevant data about date and the competition location. We created this dataset ourselves as there were only PDF versions of her competition records available online.

Below is a table containing the sources of our data:

+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| **Competition**                                          | **Link**                                                                                                |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2015 Russian Novice National Championships (Younger Age) | <https://allskaters.info/wp-content/uploads/Protocols/2014-2015/specprogramma1415_protocol.pdf>         |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2015 Tallinn Trophy (Advanced Novice)                    | <https://web.archive.org/web/20160221082023/https://data.tallinntrophy.eu/2015/Tallinn_Trophy/>         |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2016 Russian Novice National Championships (Elder Age)   | <https://allskaters.info/wp-content/uploads/Protocols/2015-2016/specprogramma1516_protocol_revised.pdf> |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2017 Russian Junior National Championships               | <https://www.fsrussia.ru/results/1617/junnat1617/index.htm>                                             |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2017 JGP Brisbane                                        | <http://www.isuresults.com/results/season1718/jgpaus2017/index.htm>                                     |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2017 JGP Minsk                                           | <http://www.isuresults.com/results/season1718/jgpblr2017/index.htm>                                     |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2017 JGP Final                                           | <http://www.isuresults.com/results/season1718/gpf1718/>                                                 |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2018 Russian Junior National Championships               | <https://www.fsrussia.ru/results/1718/junnat1718/index.htm>                                             |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2018 World Junior Championships                          | <http://www.isuresults.com/results/season1718/wjc2018/index.htm>                                        |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2018 JGP Kaunas                                          | <http://www.isuresults.com/results/season1819/jgpltu2018/>                                              |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2018 JGP Yerevan                                         | <http://www.isuresults.com/results/season1819/jgparm2018/>                                              |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2018 JGP Final                                           | <http://www.isuresults.com/results/season1819/gpf1819/>                                                 |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 Russian Senior National Championships               | <https://www.fsrussia.ru/results/1819/rusnat1819_ENG/index.htm>                                         |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 Russian Junior National Championships               | <https://www.fsrussia.ru/results/1819/junnat1819/index.htm>                                             |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 World Junior Championships                          | <http://www.isuresults.com/results/season1819/wjc2019/>                                                 |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 27th Nepala Memorial                                | <http://www.kraso.sk/wp-content/uploads/sutaze/2019_2020/20190920_mon/>                                 |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 Japan Open                                          | <https://www.jsfresults.com/InterNational/2019-2020/japanopen/>                                         |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 GP Skate Canada                                     | <http://www.isuresults.com/results/season1920/gpcan2019/>                                               |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 GP Rostlecom Cup                                    | <http://www.isuresults.com/results/season1920/gprus2019/>                                               |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2019 GP Final                                            | <http://www.isuresults.com/results/season1920/gpf1920/>                                                 |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2020 Russian Senior National Championships               | <https://www.fsrussia.ru/results/1920/rusnat1920/ENG/index.htm>                                         |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2020 European Figure Skating Championships               | <http://www.isuresults.com/results/season1920/ec2020/>                                                  |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2020 GP Rostelecom Cup                                   | <https://fsrussia.ru/results/2021/gprus2020/>                                                           |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2021 Russian Senior National Championships               | <https://fsrussia.ru/results/2021/rusnat2021/ENG/index.htm>                                             |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2021 World Championships                                 | <http://www.isuresults.com/results/season2021/wc2021/>                                                  |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2021 US International Figure Skating Classic             | <https://ijs.usfigureskating.org/leaderboard/results/2021/30212/index.asp>                              |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2021 GP Skate America                                    | <http://www.isuresults.com/results/season2122/gpusa2021/>                                               |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2022 Russian Senior National Championships               | <https://fsrussia.ru/results/2122/rusnat2122/ENG/index.htm>                                             |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2022 European Championships                              | <http://www.isuresults.com/results/season2122/ec2022/>                                                  |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+
| 2022 Winter Olympics                                     | <https://results.isu.org/results/season2122/owg2022/>                                                   |
+----------------------------------------------------------+---------------------------------------------------------------------------------------------------------+

[RELEVANT VARIABLES]{.underline}

+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Variable           | Description                                                                                                                                                                                                                                                                                                                               |
+====================+===========================================================================================================================================================================================================================================================================================================================================+
| season             | The competition season in which the jumping pass was completed. Spans two calendar years (from summer of one year to spring of the next year).                                                                                                                                                                                            |
+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| altitude_in_m      | The altitude, in meters, of the competition location.                                                                                                                                                                                                                                                                                     |
+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| distance_moscow_km | Distance, in kilometers, of the competition location from Moscow, Trusova's training city.                                                                                                                                                                                                                                                |
+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| jump               | The code entered for the jumping pass completed. The 5 types of jumps are axel (denoted by "A"), salchow ("S"), toeloop ("T"), loop ("Lo"), flip ("F"), and lutz ("Lz"). These jumps can have anywhere between 1-4 rotations and can be performed as a solo (ex. 2A) or combination jump (denoted by plus sign, ex. 3Lz+3T or 3F+2T+2Lo). |
+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| jump_bv            | The difficulty value, in points, of the jumping pass completed.                                                                                                                                                                                                                                                                           |
+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| goe_raw            | The grade of execution, a measure of the quality of the jump. Each judge - there are typically 7 or 9 - assign a GOE value between -5 (bad) to +5 (excellent). Raw GOE is calculated by dropping the highest and lowest values and finding the average of the remaining values.                                                           |
|                    |                                                                                                                                                                                                                                                                                                                                           |
|                    | Due to adjustments to the GOE system after the 2017-2018 season, GOEs for jumping passes from competitions before June 2018 are scaled up by 5/3 to adjust from the -3 to 3 scale to the -5 to 5 scale.                                                                                                                                   |
+--------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

[ETHICAL CONCERNS]{.underline}

Despite her fame within the figure skating world, Alexandra Trusova was a minor during the time period this dataset covers. As such, the only data we gathered was taken from publicly available records. We also took care to not include variables that evaluated Trusova's skating in a subjective manner and kept our analysis to quantitative data when possible.

[RESEARCH QUESTION & HYPOTHESIS]{.underline}

With this dataset, we are researching **how the difficulty of the jump and the location of the competition affect Alexandra Trusova's execution of jumps.** Through this, we seek to find patterns in Trusova's jump data that support or do not support the questions surrounding quadruple jumps in women's figure skating. We hypothesize that increases in difficulty of the jump, distance between training location and competition location, and altitude of the competition will negatively affect Trusova's execution of her jumps.

### Methodology

We plan on answering our research question using hypothesis testing and modeling.

We conducted hypothesis tests to determine whether differences in the execution score of the jump are influenced by certain variables, by comparing mean execution scores for certain predictor variable groups.

We also created linear regression models and calculated their adjusted R\^2 and AIC. We used the adjusted R\^2 and AIC as benchmarks to evaluate the effectiveness of our three variables at predicting Trusova's jump execution as they respectively penalize the addition of bad variables and penalize models with more variables.

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(tidymodels)


trusova_jump <- read.csv("data/trusova_jumps.csv")

```

[***HYPOTHESIS TESTS***]{.underline}

```{r}
#| label: jump difficulty hypothesis test

jump_difficulty <- trusova_jump |>
  arrange(jump_bv)|>
  drop_na(goe_raw) |>
  mutate(jump_group = ifelse(row_number() <= n()/2, "easy_jumps", "hard_jumps"))

difficulty_means <- jump_difficulty |>
  group_by(jump_group) |>
  summarize(mean_goe_raw = mean(goe_raw),
            n = n())

difficulty_means
```

-   **Null:** The difficulty of the jump has no significant effect on the execution of the jump.

-   **Alternate:** The difficulty of the jump affects the execution of the jump. As the difficulty of the jump increases, the execution of the jump, as defined by the raw goe, decreases.

```{r}
#| label: summary-lm difficulty v goe
summary(lm(goe_raw ~ jump_bv, data = jump_difficulty))
```

```{r}
#| label: difficulty-test plot and distribution
null_dist_jump <- jump_difficulty |>
  specify(response = goe_raw, explanatory = jump_group) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("easy_jumps", "hard_jumps"))

null_dist_jump |>
  ggplot(
    aes(x = stat)
  ) + 
  geom_histogram() + 
  geom_vline(xintercept = 0.356138) +
  geom_vline(xintercept = -0.356138)

null_dist_jump |>
  get_p_value(obs_stat = 0.356138, direction = "two sided")
```

The probability of observing our sample statistic, or something more extreme, given that the true (population) difference between the mean raw score of jumps is 0.356138, is equal to \~ 0.352.

alpha = 0.05 \< p-value 0.352. Failing to reject our null hypothesis. Weak evidence to conclude the alternative hypothesis.

```{r}
#| label: altitude hypothesis test

altitude_filter <- trusova_jump |>
  arrange(altitude_in_m)|>
  drop_na(goe_raw) |>
  mutate(alt_group = ifelse(row_number() <= n()/2, "low_altitude", "high_altitude"))

altitude_means <- altitude_filter |>
  group_by(alt_group) |>
  summarize(mean_goe_raw = mean(goe_raw),
            n = n())

altitude_means
```

-   **Null:** The altitude of the competition location has no significant effect on the execution of the jump.

-   **Alternate:** The altitude of the competition location relative to Moscow affects the execution of the jump. As the altitude of the competition location relative to Moscow increases, the execution of the jump, as defined by the raw goe, decreases.

```{r}
#| label: alt-test plot and distribution
null_dist_alt <- altitude_filter |>
  specify(response = goe_raw, explanatory = alt_group) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("low_altitude", "high_altitude"))

null_dist_alt |>
  ggplot(
    aes(x = stat)
  ) + 
  geom_histogram() + 
  geom_vline(xintercept = 0.605648) +
  geom_vline(xintercept = -0.605648)

null_dist_alt |>
  get_p_value(obs_stat = 0.605648, direction = "two sided")
```

The probability of observing our sample statistic, or something more extreme, given that the true (population) difference between the mean raw score of jumps is 0.605648, is equal to \~ 0.07.

alpha = 0.05 \< p-value 0.07. Failing to reject our null hypothesis. Weak evidence to conclude the alternative hypothesis.

```{r}
#| label: distance hypothesis test

distance_filter <- trusova_jump |>
  arrange(distance_moscow_km)|>
  drop_na(goe_raw) |>
  mutate(dist_group = ifelse(row_number() <= n()/2, "short_distance", "long_distance"))

distance_means <- distance_filter |>
  group_by(dist_group) |>
  summarize(mean_goe_raw = mean(goe_raw),
            n = n())

distance_means
```

-   **Null:** The distance of the competition location from Moscow has no significant effect on the execution of the jump.

-   **Alternate**: As the distance between Moscow and the competition location increases, the execution of the jump, as defined by the raw goe, decreases.

```{r}
#| label: dist-test plot and distribution
null_dist_dist <- distance_filter |>
  specify(response = goe_raw, explanatory = dist_group) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("short_distance", "long_distance"))

null_dist_dist |>
  ggplot(
    aes(x = stat)
  ) + 
  geom_histogram() + 
  geom_vline(xintercept = 0.603055) +
  geom_vline(xintercept = -0.603055)

null_dist_dist |>
  get_p_value(obs_stat = 0.603055, direction = "two sided")
```

The probability of observing our sample statistic, or something more extreme, given that the true (population) difference between the mean raw score of jumps is 0.603055, is equal to \~ 0.068.

alpha = 0.05 \< p-value 0.068. Failing to reject our null hypothesis. Weak evidence to conclude the alternative hypothesis.

[***Visualization***]{.underline}

```{r}
#| label: difficulty-score-comparison
trusova_jump |> 
  ggplot(aes(x = jump_bv, y = goe_raw, color = jump)) + 
  geom_point() + 
  geom_smooth() + 
  facet_wrap("season") + 
  labs(title = "Relationship Between Jump Difficulty and Score", 
       subtitle = "By Competition Season", 
       x = "Difficulty of Jump", 
       y = "Score", 
       color = "Type of Jump") + 
  theme_minimal()
```

```{r}
#| label: distance-score-comparison
trusova_jump |> 
  ggplot(aes(x = distance_moscow_km, y = goe_raw, color = jump_bv)) + 
  geom_point() + 
  geom_smooth() + 
  labs(title = "Relationship Between Distance Travelled and Score", 
       x = "Distance of Competition From Moscow", 
       y = "Score", 
       color = "Difficulty of Jump") + 
  theme_minimal()
```

```{r}
#| label: altitude-score-comparison
trusova_jump |> 
  ggplot(aes(x = altitude_in_m, y = goe_raw, color = jump_bv)) + 
  geom_point() + 
  geom_smooth() + 
  geom_vline(xintercept = 156) + 
  labs(title = "Relationship Between Altitude and Score", 
       x = "Altitude with Line Drawn at Moscow", 
       y = "Score", 
       color = "Difficulty of Jump") + 
  theme_minimal()
```

[***Predictive Models***]{.underline}

```{r}
#| label: predictive model with only difficulty

difficulty_model <- linear_reg() |>
   set_engine("lm") |>
   fit(goe_raw ~ jump_bv, data = trusova_jump) 

glance(difficulty_model)$adj.r.squared

glance(difficulty_model)$AIC

```

```{r}
#| label: predictive model with only altitude
altitude_model <- linear_reg() |>
   set_engine("lm") |>
   fit(goe_raw ~ altitude_in_m, data = trusova_jump)

glance(altitude_model)$adj.r.squared

glance(altitude_model)$AIC

```

```{r}

#| label: predictive model with only distance

distance_model <- linear_reg() |>
   set_engine("lm") |>
   fit(goe_raw ~ distance_moscow_km, data = trusova_jump)

glance(distance_model)$adj.r.squared

glance(distance_model)$AIC


```

```{r}
#| label: Altitude and Difficulty of Jump

dif_alt_model <- linear_reg() |>
  set_engine("lm") |>
  fit(goe_raw ~ jump_bv + altitude_in_m, data = trusova_jump) 

glance(dif_alt_model)$adj.r.squared
		
glance(dif_alt_model)$AIC

```

```{r}
#| label: Model with Distance and Difficulty of Jump
	
dd_model <- linear_reg() |>
  set_engine("lm") |>
  fit(goe_raw ~ jump_bv + distance_moscow_km, data = trusova_jump) 

glance(dd_model)$adj.r.squared
		
glance(dd_model)$AIC

```

```{r}
#| label: Model with Altitude and Distance

alt_dist_model <- linear_reg() |>
  set_engine("lm") |>
  fit(goe_raw ~ altitude_in_m + distance_moscow_km, data = trusova_jump) 

glance(alt_dist_model)$adj.r.squared
		
glance(alt_dist_model)$AIC

```

```{r}
#| label: Model with Altitude and Distance and Difficulty of Jump

combined_model <- linear_reg() |>
  set_engine("lm") |>
  fit(goe_raw ~ jump_bv + altitude_in_m + distance_moscow_km, data = trusova_jump) 

glance(combined_model)$adj.r.squared

glance(combined_model)$AIC

```

### Results

*The results from our hypothesis tests are as follows:*

*The results from our predictive models are as follows:*

The adjusted R squared values were less than 0.05 for all of the models. The model that predicted Trusova's raw score using only the altitude of the competition location had the highest adjusted R squared value of 0.04228525. This value, however, is still low, which indicates that none of these regression models are good fits.

The AIC values were above 1400 for all of the models, which indicates that none of the models are good fits. The model that predicted Trusova's raw score using only the altitude of the competition location had the lowest AIC value of 1438.663. 

The adjusted R squared and AIC values thus both indicate that the altitude_model is the best at predicting Trusova's raw score. However, this model is still not a good fit. There are two possibilities here: (1) that we are looking at incorrect variables, and (2) that there is no singular factor that plays a large role in a skater's performance, with critics' claims regarding jump consistency and overall execution being unfounded.

### Conclusion

Our results indicate that there is no singular factor that impacts Trusova's jump performance. We failed to reject our null hypothesis in all of our hypothesis tests, and all 7 of our predictive models had AIC values above 1430 and adjusted R squared values below 0.05.

Our data is valid because it is gathered from official competition data from the International Skating Union and individual countries' skating federations. However, it is not necessarily reliable for several reasons: 1) Due to the guidelines for GOE being rather up for interpretation, there is a small amount of bias and variability inherent in figure skating judging. This can create inconsistencies in scoring across competitions. 2) The data focuses on one skater and only her performances at national championships and international competitions. This small sample size allows for variability in patterns in the data. There are only a handful of female skaters who have performed numerous types of quads over many years of their career, so it was simply not possible to gather data for a large sample of skaters. Additionally, we had to create our dataset ourselves by combing through online PDFs of competition results, so it was not feasible to obtain data from more jumpers within the time frame of the project.

As such, the conclusions we've drawn regarding Alexandra Trusova's jumps and performance cannot be generalized to the sport of figure skating as a whole until it can be supported by data gathered from the careers of other skaters. This is mainly the result of our methods, and we could improve our analysis by gathering the same information for more female skaters that have performed quadruple jumps over their entire careers. In future research, additional data of varying types should be collected beyond what is contained in publicly available records to find other potential predictors of jump outcome. More skater-specific sensitive information, like body weight, injury status, and mental health state could also prove to create patterns in the quality of a jumping pass's execution. Because this information is not online, it would be necessary in future research to gather this data from a cohort of skaters for a duration of several seasons. Additionally, different types of hypothesis testing could be performed to potentially find a better predictor of jump outcome.

While we cannot generalize the analysis we've done of one of the most talented figure skaters in the world to every other skater, there are a few key takeaways specific to Alexandra Trusova herself. It is notable that no variables appear to be able to predict the execution of Trusova's jumps. That is, no matter the factors involved, Trusova performs her jumps to a similar level of quality every time. This consistency speaks to Trusova's high level of training. It also speaks to how even Olympic-level figure skaters fall and mess up periodically with no particular pattern.
